Tutorial: SQL-on-Hadoop Systems

### 2015-04-01 paper.tex
Adding BigSQL, Impala, and refs

\numberofauthors{1}

% 1st author
\author{
}

@@

An important aspect of SQL-on-Hadoop, which is orthogonal to all these categories, is the various HDFS file formats used to store the data. While text is an important format for in-situi SQL analysis, and ELT processing, various columnar formats have been proposed to provide fast analytics query processing. In this tutorial, we will first introduce these file formats, then we propose to describe the SQL-on-Hadoop systems in detail, examine their architectural differences, and cover in-depth their approach to indexing, updates, query optimization, as well as new algorithms they introduce in a 3-hour tutorial.

@@

@@

Spark is a fast general purpose cluster computing engine that is compatible with Hadoop data.  There are three different systems that use Spark as their run-time for SQL processing: Shark \cite{sharksigmod13}, Hive on Spark \cite{hiveOnSpark}, and Spark SQL \cite{sparkSQL}. Shark replaces the MapReduce run-time of Hive \cite{hive} with Spark. It is still uses the Hive compiler, and interfaces but genereates Spark jobs. Hive on Spark is another attempt at replacing MapReduce/Tez based run-times of Hive with Spark. Spark SQL uses the Shark \cite{sharksigmod13} run-time but replaces the HiveQL and compiler with SQL, and a new compiler that uses the catalyst query optimizer.

>>>>

\numberofauthors{3}

% 1st author
\author{
\alignauthor
Daniel Abadi\\
\affaddr{Yale University}\\
\email{daniel.abadi@yale.edu}
\alignauthor
Fatma  {\"O}zcan\\
\affaddr{IBM}\\
\email{fozcan@us.ibm.com}
\alignauthor
Ippokratis Pandis\\
\affaddr{Cloudera}\\
\email{ippokratis@cloudera.com}
}

@@

An important aspect of SQL-on-Hadoop, which is orthogonal to all these categories, is the various HDFS file formats used to store the data. While text is an important format for in-situ SQL analysis, and ELT processing, various columnar formats have been proposed to provide fast analytics query processing. In this tutorial, we will first introduce these file formats, then we propose to describe the SQL-on-Hadoop systems in detail, examine their architectural differences, and cover in-depth their approach to indexing, updates, query optimization, as well as new algorithms they introduce in a 3-hour tutorial.

@@

Cloudera Impala \cite{Kornacker+15} an open-source, fully-integrated MPP SQL query engine. Impala's goal is to combine the familiar SQL support and multi-user performance of a traditional analytic database with the scalability and flexibility of Apache Hadoop. Unlike other systems (often forks of Postgres), Impala is a brand-new engine. It is written from the ground up in C++ and Java, and employs uses LLVM to generate code at runtime in order to speed up frequently executed code paths. It maintains Hadoop's flexibility by utilizing standard components (HDFS, HBase, Metastore, YARN, Sentry) and is able to read the majority of the widely-used file formats (e.g. Parquet, Avro, RCFile). To reduce latency, such as that incurred from utilizing MapReduce or by reading data remotely, Impala implements a distributed architecture based on daemon processes that are responsible for all aspects of query execution and that run on the same machines as the rest of the Hadoop infrastructure. The result is performance that is on par or exceeds that of commercial MPP analytic DBMSs, depending on the particular workload.

IBM Big SQL \cite{bigsql} leverages IBM's state-of-the-art relational database technology, without imposing any database structures or restrictions, and processes standard SQL queries over HDFS data, supporting all common Hadoop file formats, including text and sequence files, as well as Parquet and ORC files. Big SQL does not introduce any propriety data format. Big SQL 3.0 shares the same catalog and table definitions with Hive using the Hive Metastore. Database workers read HDFS data directly and execute relational operations. Big SQL exploits sophisticated query rewrite transformations \cite{pirahesh96, winmagic} that are targeted for complex nested decision support queries. It uses sophisticated data statistics and a cost-based optimizer to choose the best query execution plan.  In traditional shared-nothing databases, each worker owns a partition of the data. Big SQL deviates from this basic model and introduces a scheduler service that assigns HDFS blocks to database workers for processing on a query by query basis. The scheduler identifies where the HDFS blocks are and decides which database workers to include in the query plan. The assignment is done dynamically at run-time to accommodate failures: scheduler uses the workers that are currently up and running. This also allows elasticity. If a new node is added to the database cluster, it can be considered immediately by the scheduler for the new queries. Similarly, if a node crashes or the cluster is scaled down, the scheduler immediately detects this change and chooses database workers for future queries accordingly.

@@

Spark is a fast general purpose cluster computing engine that is compatible with Hadoop data.  There are three different systems that use Spark as their run-time for SQL processing: Shark \cite{sharksigmod13}, Hive on Spark \cite{hiveOnSpark}, and Spark SQL \cite{sparkSQL}. Shark replaces the MapReduce run-time of Hive \cite{hive} with Spark. It is still uses the Hive compiler, and interfaces but generates Spark jobs. Hive on Spark is another attempt at replacing MapReduce/Tez based run-times of Hive with Spark. Spark SQL uses the Shark \cite{sharksigmod13} run-time but replaces the HiveQL and compiler with SQL, and a new compiler that uses the catalyst query optimizer.
